Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils

### 2016-02-08 manuscript/cpcrw_manuscript.Rmd
Update cpcrw_manuscript.Rmd
added methods text

In the lab, the soil cores were stored at 4 °C for several days until they were weighed and prepared for incubation. At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group). These groups included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume. The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 °C "drought" group. The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density. Subsamples were collected and stored at -20 °C for potential enzyme activity and ATP analysis.

@@

For each measurement, a six-core treatment group was connected to a Picarro A0311 multiplexer that was in turn connected to a Picarro G2301 GHG Analyzer (Picarro Inc., Santa Clara, CA, USA). Dry CH~4~ and CO~2~ concentrations were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group. Cores were weighed immediately after (PEYTON) gas measurements. Ambient air was measured between treatment groups, and before starting measurements in a chamber.

The incubation experiment concluded on 10 December 2015. PEYTON: text re C, N, etc. analyses.

>>>>

In the lab, the soil cores were stored at 4 °C for several days until they were weighed and prepared for incubation. At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group). These groups included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume. The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 °C "drought" group. The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density. Subsamples were collected and stored at -20 °C for dissolved organic carbon measurements.

@@

For each measurement, a six-core treatment group was connected to a Picarro A0311 multiplexer that was in turn connected to a Picarro G2301 GHG Analyzer (Picarro Inc., Santa Clara, CA, USA). Dry CH~4~ and CO~2~ concentrations were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group. Cores were weighed immediately after gas measurements. Ambient air was measured between treatment groups, and before starting measurements in a chamber. BEN: How was ambient air measurements used in this study? don't see any mention of it in remaining text, etc.

The incubation experiment concluded on 9 December 2015, following the final CO~2~ and CH~4~ readings. Each soil core was maintained at the treatment-dependent temperature and moisture content (by weight) until removed for destructive sampling, December 14 - 18, 2015. Subsamples were collected and composited throughout each soil core for future dissolved organic carbon analysis (110 +/- 24 g dry weight equivalent) and dry-weight calculations (~ 28 g) upon destrucitve core sampling. The remaining core material was air-dried and then separated into particles > 2mm diameter and soil < 2mm diameter using a U.S. Standard Test Sieve, No. 10 (Fisherbrand,  Pittsburg, PA, USA). Both the dry weight and the volume of soil ( < 2 mm) were used in calculations of gravimetric and volumetric soil moisture content (_Gardner, WH. 1986 Chapter 21 in Klute, A., Ed. Methods of Soil Analysis. Part 1. Physical and mineralogical methods. 2nd ed. Soil Science Society of American, Inc. Madison, Wisconsin_). Soil volume was caclulated as the total core volume - volume of particles greater than 2 mm diameter, determined using water displacement of particles > 2 mm diameter. Air-dried soil (< 2 mm diamter) and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia extension in February, 2016 for total C, N of dried soil samples and dissolved organic carbon from samples stored at -20 °C.
